---
title: 'Homework 1: Basic Data Wrangling '
author: "Philipp Masur"
date: ""
output:
  html_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r opts, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, results = TRUE, message = FALSE, warning = FALSE)
```

# Formalities

- Name:         [ENTER YOUR NAME HERE]
- Student ID:   [ENTER YOUR STUDENT ID HERE]   

In the end, klick on "knit" and upload the respective html-output file to Canvas. Please add your name and lastname to the output file name:
e.g., 01_homework_assignment_NAME-LASTNAME.html

# Instructions

In this homework assignment, you are going to practice what you have learned in the past two practical sessions. To do soe, we need to get some data into R. On Canvas, you will find a module called "Data Sets". Within this module, there are several data sets that we will use throughout the course. For this assignment, download the data set **norm_data.csv**. This data set stems from an online survey on social media use and norm perceptions. 

If you have set the working directory to the location of this RMarkdown Script and you put the csv-file in the same folder, you can load that data set without specifying a specific path. 

```{r}
library(tidyverse)
d <- read_csv("norm_data.csv")
head(d)
```


# Exercises

## Task 1: Provide a description of the sample

Let's explore the data set. The data set contains some variables that assess socio-demographic characteristics of the participants (age, gender, edu, race...). Use the tidyverse functions that we discussed on the last practical session, 

- Create a subset that only contains socio-demographic variables.
- How old are participants on average?
- How many female and male participants are in the data set (0 = male, 1 = female, 2 = other)?
- How old are participants from different educational levels?

```{r}
# Solution here
```

## Task 2: Explore social media use 

The following code adds a mean indices that represent the latent dimension *critical media literacy* (ranging from 1 to 7). 

- Create a subset that contains the variables *age*, *edu*, and *MLI_CRI*.
- Remove all missing values using the function `na.omit()`.
- Does critical media literacy differ depending on the educational level?
- Filter out people below 40 and again assess whether critical media literacy differs across educational levels.

```{r}
d <- d %>%
  mutate(MLI_CRI = rowMeans(d %>% select(contains("MLI_CRI")), na.rm = T))

# Solution here
```


