---
title: 'Homework 1: Basic Data Wrangling (Solutions) '
author: "Philipp Masur"
date: ""
output:
  github_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r opts, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, results = TRUE, message = FALSE, warning = FALSE)
```

# Formalities

- Name:         [ENTER YOUR NAME HERE]
- Student ID:   [ENTER YOUR STUDENT ID HERE]   

In the end, klick on "knit" and upload the respective html-output file to Canvas. Please add your name and lastname to the output file name:
e.g., 01_homework_assignment_NAME-LASTNAME.html

# Instructions

In this homework assignment, you are going to practice what you have learned in the past two practical sessions. To do soe, we need to get some data into R. On Canvas, you will find a module called "Data Sets". Within this module, there are several data sets that we will use throughout the course. For this assignment, download the data set **norm_data.csv**. This data set stems from an online survey on social media use and norm perceptions. 

If you have set the working directory to the location of this RMarkdown Script and you put the csv-file in the same folder, you can load that data set without specifying a specific path. 

```{r}
library(tidyverse)
d <- read_csv("norm_data.csv")
head(d)
```


# Exercises

## Task 1: Provide a description of the sample

Let's explore the data set. The data set contains some variables that assess socio-demographic characteristics of the participants (age, gender, edu, race...). Use the tidyverse functions that we discussed on the last practical session, 

- Create a subset that only contains socio-demographic variables.
- How old are participants on average?
- How many female and male participants are in the data set (0 = male, 1 = female, 2 = other)?
- How old are participants from different educational levels?

```{r}
# Create subset
d_soc <- d %>%   ## don't forget to save the subset to be able to work with it afterwards!
  select(age, gender, edu, race, employment, relationship)

# Recode one person who provided birth year instead of age (This is a good reminder to always check your data!)
d_soc <- d_soc %>%
  mutate(age = ifelse(age > 100, 2021-age, age))

# Alternative 1: Using base functions
mean(d_soc$age, na.rm = T)
sd(d_soc$age, na.rm = T)

# Alternative 2: Using the summarize function in the tidyvere
d_soc %>%
  summarize(m = mean(age, na.rm = T),
            sd = sd(age, na.rm = T))

# Alternative 3: Using the psych package and its function describe
library(psych)  # you have to install this first!
describe(d_soc$age) 
```

**Answer:** Participants are on average M = 36.5 years old (SD = 11.3). 

```{r}
# Alternative 1: Frequency table of gender using summarize
d_soc %>%
  group_by(gender) %>%
  summarize(n = n()) %>%
  mutate(prop = prop.table(n))  ## adding percentages to the table

# Alternative 2: Using the count function
d_soc %>%
  group_by(gender) %>%
  count %>%                     ## Does the same as the summarizing command above
  mutate(prop = prop.table(n))  
```

**Answer:** There are 411 males (35.9%) and 638 females (55.7%) in the sample. Five participants did not indicate their gender and 92 did not indicate their gender at all.

```{r}
# Age per educational level
d_soc %>%
  group_by(edu) %>%
  summarize(m = mean(age, na.rm = T))
```

**Answer:** Age hardly differs between educational levels (range = 34.2 - 37.9). 

## Task 2: Explore social media use 

The following code adds a mean indices that represent the latent dimension *critical media literacy* (ranging from 1 to 7). 

- Create a subset that contains the variables *age*, *edu*, and *MLI_CRI*.
- Remove all missing values using the function `na.omit()`.
- Does critical media literacy differ depending on the educational level?
- Filter out people below 40 and again assess whether critical media literacy differs across educational levels.

```{r}
d <- d %>%
  mutate(MLI_CRI = rowMeans(d %>% select(contains("MLI_CRI")), na.rm = T))

# Select variabke, remove all missings and create the subset
d2 <- d %>%                        ## again, don't forget to save your steps in a new object!
  select(age, edu, MLI_CRI) %>%
  na.omit

# Summarize critical media literacy by educational level
d2 %>%
  group_by(edu) %>%
  summarise(mean = mean(MLI_CRI, na.rm = T),
            sd = sd(MLI_CRI, na.rm = T))

# Bonus: Significant test using an analysis of variance (ANOVA)
m <- aov(MLI_CRI ~ edu, d2)
summary(m)

```

**Answer:** Based on a simple mean comparison, critical media literacy does differ between educational level: The higher the educational level, the more literate are the participants. This difference is significant, F(1, 1051) = 13.96, p < .001. 

```{r}
# Create subset
d2_old <- d2 %>%
  filter(age >= 40)

# Summarize data in subset
d2_old %>%
  group_by(edu) %>%
  summarize(mean = mean(MLI_CRI, na.rm = T),
            sd = sd(MLI_CRI, na.rm = T))

# Bonus: Significant test using an analysis of variance (ANOVA)
m <- aov(MLI_CRI ~ edu, d2_old)
summary(m)
```

**Answer:** If we only inspect older participants (>= 40 years), this results changes slightly: the differences are seemingly smaller. Yet, the difference is still statistically significant: F(1,334) = 4.65, p = .032. 

## Task 3: Norm perceptions on Instagram

The data set contains some variables that measure norm perceptions in relation to visual self-disclosure on Instagram (12 items starting with "INV_NOR_"). 

- Create the mean across all these item per person and add it as a new column to the data set.
- Do norm perceptions differ depending on gender (only looking at 0 = males and 1 = females)?
- How could you test whether this difference is statistically significant (tip: ?t.test)? 


```{r}
d <- d %>%
  mutate(NORM = rowMeans(d %>% select(contains("INV_NOR")), na.rm = T))

# Filter 
d_gen <- d %>%
  filter(gender == 0 | gender == 1)

# Means
d_gen %>%
  group_by(gender) %>%
  summarize(average_norm = mean(NORM))

# Significant test
t.test(NORM ~ gender, d_gen)

```

**Answer:** Males perceive norms related to visual self-disclosure to be slightly less strong (M = 4.90) compared to females (M = 5.1). This difference is statistically significant, t(846.76) = -4.00, p < .001. 


