---
title: 'Homework 2: Data Wrangling and Data Visualization'
author: "Philipp Masur"
date: ""
output:
  html_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r opts, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, results = TRUE, message = FALSE, warning = FALSE)
```

# Formalities

- Name:         [ENTER YOUR NAME HERE]
- Student ID:   [ENTER YOUR STUDENT ID HERE]   

In the end, klick on "knit" and upload the respective html-output file to Canvas. Please add your name and lastname to the output file name:
e.g., 02_homework_assignment_NAME-LASTNAME.html

# Instructions

In this homework assignment, you are again going to practice what you have learned in the past two practical sessions. To do so, we again need to get some data into R. In the Canvas Module "Homework Templates & Data Sets" you will find the data set **ESS9_2018.csv**. This data set is the most recent European Social Survey (Wave 9). It can be downloaded from this website: http://www.europeansocialsurvey.org/data/, where you can find also more information about the data set. 

If you have set the working directory to the location of this RMarkdown Script and you put the csv-file in the same folder, you can load that data set without specifying a specific path. 

```{r}
library(tidyverse)
library(ggthemes)
d <- read_csv("ESS9_2018.csv")
```


# Exercises

## Task 1: Selecting the right variables and understanding the data

1.1. The data set is comparatively big and we only want to work with some variables. So in a first step, select and potentially rename the following variables:

- **idno**:     respondents' id
- **cntry**:    country
- **nwspol**:   News consumptions (in minutes per day) about politics and current affairs (watching, reading, listening)
- **netustm**:  Internet use (in minutes per day)
- **polintr**:  Political interest (1 = very interested, 2 = quite interested, 3 = hardly interested, 4 = not at all interested)
- **stflife**:  Life satisfaction 
- **gndr**:     Gender
- **agea**:     Age (in years)


```{r}
d <- d %>%
  select(id = idno,
         country = cntry,
         pol_news = nwspol,
         internet_use = netustm,
         pol_interest = polintr,
         trust_parties = trstprt,
         life_sat = stflife,
         pol_ideology = prtdgcl,
         gender = gndr,
         age = agea)
```

2.2. How many people are in the data set and what is their average age? How many people are there in each country?

```{r}
nrow(d)
summarise(d, age = mean(age, na.rm = T))

d %>%
  group_by(country) %>%
  summarize(n = n()) %>%
  arrange(-n)
```

**Answer:** There are 49,297 people in the data set. They are on average 51.1 years old. The number of people per country differs, ranging from 781 (Cyprus) to 2745 people (Italy).


## Task 2: Summarizing across countries

2.1. Try to figure out on what scale political interest was measured. What is the mean and standard deviation of the variables "political interest", "political news consumption"? Tip: Install the package `psych` and use the function `describe()` to assess the variables psychometric properties. 

```{r}
# install.packages("psych")
library(psych)
d %>%
  select(pol_interest,
         pol_news) %>%
  describe
```

**Answer:** The scale measuring political interest was measure on a 4-point-scale. On, average, participants were mildly interested in politics (M = 2.66, SD = 0.92). On average, they watched, read, or listen to political news M = 85.48 minutes per day (SD = 137.59)

2.2. Check the assumption that both variables are normally distributed. Do so by plotting histograms for both variables.

```{r}
# Political interest
ggplot(d, aes(x = pol_interest)) +
  geom_histogram(color = "white", fill = "lightblue", bins = 4) +
  theme_clean()

# Political news consumption
ggplot(d, aes(x = pol_news)) +
  geom_histogram(color = "white", fill = "lightblue", bins = 50) +
  theme_clean()
```

**Answer:** As political interest was only measured on a 4-point scale, it is difficult to assess whether it is normally distributed but the histogram indeed suggest that the values 2 and 3 are more common the extremes. News consumption on the other hand, is clearly not normally distributed. Many people seem to watch, read, or listen to political news between 0 and 100 minutes per day. The distribution, however, has a quite long tail. So there are a few participants in the survey that consume much more news per day. 

2.2. Second, try to use the `group_by` and `summarise` commands to produce the average news consumptions across countries. If you can, also estimate the standard deviation per country.

```{r}
news_per_country <- d %>%
  group_by(country) %>%
  summarise(m = mean(pol_news, na.rm = T),
            sd = sd(pol_news, na.rm = T))
news_per_country
```

2.3. If you have summarized the data, assign them to a new object and try to plot difference in news consumption between countries as a barplot. Reorder the countries to show which is consumes most political news in which the least. Which countries consume the most political news?

```{r}
library(ggthemes)
news_per_country %>%
  ggplot(aes(x = fct_reorder(country, m), y = m)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_clean() +
  labs(x = " ", y = "Political news consumption (Minutes/day)")

```

**Answer:** Overall, average news consumption varies considerably across countries(ranging from 45 minutes in Czech Republics to >150 minutes per day in Italy). People in Belgium and Italy consume most political news per day. 


## Task 3: Does internet use affect well-being?

The subsequent code runs a linear model in which life satisfaction is predicted by internet use. Some further variables are entered as control variables. Based on the results of this model, Internet use is a positive, significant predictor of life satisfaction. 

3.1. How could you supplement this analysis with an appropriate graph? Is the relationship really practically relevant?

```{r}
# Linear model
model <- lm(life_sat ~ age + gender-1 + internet_use, d)
summary(model)

# Scatterplot
ggplot(d, aes(x = internet_use, y = jitter(life_sat))) +
  geom_point(alpha = .2, color = "black", size = 1) +
  geom_smooth(method = "lm") +
  theme_clean() +
  labs(x = "Internet use (minutes/day", y = "Life satisfaction")

```


3.2. Does the relationship differ for different age groups? The following code creates a new age variable. Add this variable age_group to the aesthetics of the plot you created above (e.g., by adding `color = age_group?`) What do you see? 

```{r}
# Creating a age variable, filtering out people who did not provide age
age_data <- d %>%
  mutate(age_group = cut(age, 3)) %>%
  filter(!is.na(age_group))

# Plot relationships across the three groups
ggplot(age_data, aes(x = internet_use, y = jitter(life_sat), color = age_group)) +
  geom_smooth(method = "lm") +
  theme_clean() +
  labs(x = "Internet use (minutes/day", y = "Life satisfaction")
```

**Answer:** Whereas internet use seems to slightly increase life satisfaction for younger generation, the relationship is clearly negative for people above 65 (blue line).


3.3. Bonus questions: Can you add a simple command to the plot that creates the same type of visualization for all countries? What do you see?

```{r, fig.width = 10, fig.height = 10}
ggplot(age_data, aes(x = internet_use, y = jitter(life_sat), color = age_group)) +
  geom_smooth(method = "lm") +
  facet_wrap(~country) +
  theme_bw() +
  labs(x = "Internet use (minutes/day", y = "Life satisfaction", color = "Age groups")
```

**Answer:** The relationships for the three age groups vary considerably across the countries. For example, in Portugal (PT), the relationship between internet use and life satisfaction is clearly negative for older people, but this is not necessarily true for other countries (e.g., Italy). 


