---
title: 'Homework 4: Advanced Supervised and Unsupervised Text Classification'
author: "Philipp Masur"
date: ""
output:
  html_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r opts, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, results = TRUE, message = FALSE, warning = FALSE)
```

# Formalities

- Name:         [ENTER YOUR NAME HERE]
- Student ID:   [ENTER YOUR STUDENT ID HERE]   

In the end, klick on "knit" and upload the respective html-output file to Canvas. Please add your name and lastname to the output file name:
e.g., 02_homework_assignment_NAME-LASTNAME.html


# Introduction

In this homework assignment, you are going to practice several things that we have learned throughout the course. So in order to solve and answer the questions, you may have to go back to several handouts and not just the ones from this week. 

## Loading packages and data

For this assignment, you are going to work with a data set that contains app reviews for TikTok from the google play store. Each review comes with upvotes. We are going to subset the dataset to only the first 5000 reviews, so that it won't take as long to fit different models. 

```{r}
library(tidyverse)
library(quanteda)
library(quanteda.textplots)
library(quanteda.textmodels)
library(topicmodels)
library(caret)


d <- read_csv("tiktok_reviews.csv") %>%
  select(text = review_text, upvotes, app_version, posted_date) %>%
  mutate(cat_votes = case_when(upvotes == 0 ~ "none",
                               upvotes > 0 & upvotes < 20 ~ "few",
                               upvotes >= 20 ~ "many"))
head(d)
```


## Explore the data set

**Question:** What are the most frequently used words in the corpus? (Tip: You may want to remove some very common words such as e.g., "tiktok"). Can you plot this?

```{r}
# Solution here
```

**Answer:** [write your answer/interpretation here]

# Supervised machine learning

**Question:** Use machine learning models to predict the upvotes (use the categrized version cat_votes) from the app reviews. Test both Naive Bayes and Support Vector Machines algorithms. Which performs better? What can you learn from the confusion matrix and respective performance scores? Where does the algorithm perform particularly bad?

```{r}
# Solution here
```

**Answer:** [write your answer/interpretation here]


# Unsupervised text classification

**Questions:** What topics can you find in the app reviews? Use topic modeling to meaningfully classify the data set. I already did some text preprocessing as this is somewhat difficult here. Pay attention to the last line: here, I am filtering out words that are too frequent and thus don't really help in creating meaningful topics. Now, try out models with different numbers of topics and make sense of the output. Decide which one works best (you may want to take perplexity into account). Best on your chosen solutions, which topics are most often occurring (can you plot this?)? 

```{r}
# Solution here
```


**Answer:** [write your answer/interpretation here]
