---
title: 'Visualizing data with ggplot'
author: "Wouter van Atteveldt"
output: 
  github_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE, remove_for_md=TRUE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```



## Introduction

Creating visualizations is one of the most useful and powerful ways to analyse and data
and tell a convincing story with these data. 

In `ggplot`, R Tidyverse has a very powerful and well-designed visualization library.
It can be used to make almost all kinds of graphs, 
from simple line and scatter plots to colorful annotated bubble plots. 

For an an overview of possibilities and more explanation, you can check out the free online resources below:

- [The R Graph gallery](https://r-graph-gallery.com/ggplot2-package.html) - contains many examples of beautiful plots, all with R code included
- [From data to viz](https://www.data-to-viz.com/) - an interactive site that showcases good plots based on your data, giving not just the R code but also best practices, common mistakes, and alternatives for each plot. 
- [The ggplot cheat sheet](https://github.com/rstudio/cheatsheets/blob/main/data-visualization.pdf) with a nice overview of ggplot functions and options
- [Data Visualization](https://socviz.co/) by Kieran Healy - a completely free online book that discusses the principles of data visualization and the  best ways to visualize various types of data, also including R code for all examples.

::: Info ::::::

Most of the plots in this tutorial focus on the mechanics of the visualization and don't spend a lot of attention on adding informative labels, making it visually attractive, etc. 
For plots in reports and presentations, however, it is extremely important to pay attention to these details.
So, be sure to check out the last section on *Creating high-quality graphs* as well!
:::

### GGplot basics

To illustrate how ggplot works, let's take a look at a simple built-in data set of car tests:

```{r}
library(tidyverse)
data(mtcars)
mtcars
```


We can visualize the relation between engine power and fuel economy:

```{r basicplot}
ggplot(mtcars, 
       aes(x=hp, y=mpg)) + 
  geom_point()
```

**Solutions:**

```{r basicplot-solution}
ggplot(mtcars, 
       aes(x=wt, y=mpg)) + 
  geom_point()
```


**Exercise:** First, let's run the code the way it is now (press the 'run code' button). Now, can you change the plot to show the `wt` (weight) on the x axis instead?

### Elements of a ggplot

The ggplot call above is typical for how ggplot works,
and showcases the *grammar of graphics*. Essentially, every graph contains these three basic elements:

- The **data**, in this case `mtcars`. This is the first argument of the ggplot function.
- One or more **geometrical elements** or `geom`s, that form the actual visualization. 
- An **aesthetic mapping** or `aes` that maps columns in the data to aspects of the geometry.

In other words, the plot above is a plot of points (i.e. a scatter plot),
whose x and y positions are mapped to the `hp` and `mpg` columns of the data frame `mtcars`

### Make it pretty!

Besides these three basic elements (data, geoms, and aesthetics),
you can tweak almost every aspect of a plot. 
For example, the code below adds an extra mapping, making the colors of points reflect the number of cylinders of the car. 

Besides this, it adds three extra elements:

- A `scale`. By default, ggplot looks at the data and determines a reasonable scale. This default is generally quite good,
  and means that you can quickly make decent graphs. With the `scale_*` commands, you can tweak this scale, for example by 
  changing color values, the range, etc.
- A `ggtitle`. You can add various elements like titles, subtitles, axis labels etc. to the plot to make it more informative
- A `theme`. Themes determine the look and feel of everything around the data, e.g. the plot background, grid lines, fonts, etc. 
  You can create and tweak themes to change the look and feel of plots, for example to match the house style of a company or journal. 

```{r prettyplot}
ggplot(mtcars, aes(x=hp, y=mpg, color=cyl)) + 
  geom_point() + 
  scale_color_gradient(low="darkgreen", high="darkred") +
  ggtitle("Car horsepower and fuel economy") + 
  theme_minimal()
```

Feel free to play around, e.g. by inverting the x and y in the mapping, changing the colors, adding axis labels (`+ xlab("Horsepower")`), etc. 

## Scatter plots

A very common but quite insightful plot is the scatter plot. It is useful if you want to express the relation between two numeric variables. 

As you saw above, scatter plots use the `geom_point()` geometry. For this plot type, you have to specify and **x** and **y** mapping. 
In addition, you can specify a `color`, `fill`, and `alpha` (transparency) mapping, as well as `size` and `shape`.

For example, the plot below shows the relation between population density and percentage of people aged 65+ in Dutch municipalities.
It also uses point size to indicate total population size, and color to indicate income 
(see e.g. [http://sape.inf.usi.ch/quick-reference/ggplot2/colour](http://sape.inf.usi.ch/quick-reference/ggplot2/colour)) for a reference of color names in ggplot)

```{r point}
library(tidyverse)
demographics <- read_csv('https://raw.githubusercontent.com/vanatteveldt/ccslearnr/master/data/dutch_demographics.csv')
ggplot(demographics,
       aes(x = v57_density, y = c_65plus, size=v01_pop, color=v132_income)) +
  scale_color_gradient(low="coral", high="darkblue") + 
  geom_point(alpha=.6)
```

**Solution:**

```{r point-solution}
library(tidyverse)
demographics <- read_csv('https://raw.githubusercontent.com/vanatteveldt/ccslearnr/master/data/dutch_demographics.csv')
ggplot(demographics,
       aes(x = v57_density, y = c_65plus, size=v01_pop, color=v43_nl)) +
  scale_color_gradient(low="darkblue", high="lightblue") + 
  geom_point(alpha=.6)
```



**Exercise**: Can you change the code to make the colors represent the percentage of inhabitants without a migration history (`v43_nl`), 
and change the color scale to run from dark blue to light blue?s


Note the use of `alpha` to make the points somewhat transparent, making it a bit easier to see when points overlap. 
This is given as a constant value rather than a mapping, since we want all points to have this.
This can be used on all geoms: if you wish to e.g. make all points a certain color or size,
you add these in the `geom(...)` call directly, i.e. not in an `aes(...)` call.

If you want to play around a bit more: What happens when you add `shape=21` or even `shape='ðŸ '` to the call? 
For `shape=21`, you can also add a fill aesthetic (or constant value) if desired - how does that affect the plot? 
(you can google 'ggplot shapes' to see an overview of built-in default shapes, like the circles for 21)


## Line graphs

Line graphs mostly useful to showcase change over time. 
This uses `geom_line`, which is in many ways similar to `geom_point`, but rather than drawing points, it will draw a line between all points.
Besides the `x`, `y`, `color` and `alpha` mapping, it can have a `linetype` (e.g. dashed, dotted, solid) and a `linewidth`.

For example, the following code shows the proportion of the popular vote for each US Presidential candidate since 2000:

```{r line}
library(tidyverse)
elections <- read_csv('https://raw.githubusercontent.com/vanatteveldt/ccslearnr/master/data/US_Elections_years.csv')
ggplot(elections,
       aes(x=year, y=votes, color=party)) + 
  geom_line()
```

### Let's make it pretty!

To make this graph a bit nicer, let's make some changes:
- The votes are given as a proportion of 0-1, but most people prefer 0-100%. This can be changed in the `scale` by setting a labelling function. 
- The party names are in all capitals, which is ugly and distracting. We can use `str_to_title` to change them to Title Case. 
- This graph uses the default colors for parties, which do not match our normal ideas of which party should have which color.
  We can fix this by using a manual scale for the color: `scale_color_manual`, which allows us to specify exactly which value should get which color. 
- The data for the minor parties was sometimes missing. For example, the Greens were only included in 2000 and 2020. 
  By default, ggplot draws a line between those points, which misleadingly suggests that they had a result in the intervening years as well.
  We fix this using the `complete` function, which inserts missing combinations of `year` and `party`, setting the `vote` variable to zero.
- Finally, let's make the line for the minor parties dotted, using `mutate` to define major, and then using it as the linetype aesthetic, 
  again setting a manual scale to force the FALSE values to be dashed (and removing the guide/legend).

To achieve these two changes, we need to preprocess the data before the `ggplot` command, using the `complete` and `mutate` functions.
We do this by placing these in a pipeline before `ggplot`, and then of course omitting the data argument from the ggplot call.
(see the handout on summarizing data for an explanation of pipelines, if needed). 

```{r electiondata}
library(tidyverse)
elections <- read_csv('https://raw.githubusercontent.com/vanatteveldt/ccslearnr/master/data/US_Elections_years.csv')
```


```{r line2}
elections |> 
  complete(year, party, fill=list(votes=0)) |>
  mutate(major = party == "DEMOCRAT" | party == "REPUBLICAN",
         party = str_to_title(party)) |>
  ggplot(aes(x=year, y=votes, color=party, linetype=major)) + 
  scale_y_continuous(labels = scales::percent) + 
  scale_linetype_manual(values=c("dashed", "solid"), guide='none') + 
  scale_color_manual(breaks=c("Democrat", "Republican", "Green", "Libertarian", "Other"),
                     values=c("blue", "red", "green4", "yellow4", "grey")) + 
  geom_line()
```

## Bar plots

Where scatter and line plots generally focus on plotting individual cases, bar plots often present some sort of summary per case.
For example, let's consider that US Demographics data:

```{r usdemo}
library(tidyverse)
us_demographics <- read_csv('https://raw.githubusercontent.com/vanatteveldt/ccslearnr/master/data/US_Demographics.csv')
us_demographics
```


Suppose we would like to show the number of counties per state. This can be done using a bar plot, for which you only specify
either the `y` or `x` axis. This by default *counts* the number of cases and creates a bar for each case representing this count:

```{r bar}
ggplot(us_demographics, aes(y=state)) + geom_bar()
```

### Histograms

For continuous variables, you need to divide the range into *bins*. This can be done using `geom_histogram`, which also takes only an `x` or `y` argument.
For example, the code below shows the number of counties by percentage non-white population, which is divided into 10 bins (deciles):


```{r hist}
ggplot(us_demographics, aes(x=nonwhite_pct)) + 
  geom_histogram(bins=10, color="black")
```

### Using geom_col instead of geom_bar

There are some more interesting options to explore with a bar plot, as showcased by the plot below:

- You can also specify the summary value manually if you have already computed the statistic, for example for the votes in the election data loaded above.
  For this, use `geom_col` rather than `geom_bar` and specify both `x` and `y` aesthetic
- To create a grouped bar plot, specify a `fill` aesthetic and add `position_dodge2()` to the geom. 
  (note: normally `position_dodge()` also works, but with a continuous grouping value like year you need `dodge2`)
- Often, a vertically arranged bar plot can be more readable than a horizontal one.
  However, if you prefer a horizontal plot, you can put the x markers at a 45 or 90 degree angle. 

```{r col}
library(tidyverse)
elections <- read_csv('https://raw.githubusercontent.com/vanatteveldt/ccslearnr/master/data/US_Elections_years.csv')
elections |>
  complete(year, party, fill=list(votes=0)) |>
  ggplot(aes(x=party, y=votes, fill=year)) + geom_col(position=position_dodge2()) + 
  scale_x_discrete(guide = guide_axis(angle = 45)) + 
  ylab("")
```

**Solution:**

```{r col-solution}
# Option 1: You can add the as.factor(year) directly into the aesthetic : ----
library(tidyverse)
elections <- read_csv('https://raw.githubusercontent.com/vanatteveldt/ccslearnr/master/data/US_Elections_years.csv')
elections |>
  complete(year, party, fill=list(votes=0)) |>
  ggplot(aes(x=party, y=votes, fill=as.factor(year))) + geom_col(position=position_dodge2()) + 
  scale_x_discrete(guide = guide_axis(angle = 45)) + 
  ylab("")

# Option 2: You can also call mutate first : ----
library(tidyverse)
elections <- read_csv('https://raw.githubusercontent.com/vanatteveldt/ccslearnr/master/data/US_Elections_years.csv')
elections |>
  complete(year, party, fill=list(votes=0)) |>
  mutate(year=as.factor(year)) |>
  ggplot(aes(x=party, y=votes, fill=year)) + geom_col(position=position_dodge2()) + 
  scale_x_discrete(guide = guide_axis(angle = 45)) + 
  ylab("")
```



(note the use of complete to ensure that missing year-party combinations are treated as zero).

**Exercise:** In the code above, `year` is a numeric variable, causing R to create a gradient scale for it.
Can you change year into a factor (using `as.factor(year)`) in the code above? 
How does that change the plot?

## Combining geoms

The plots below all had a single `geom`, creating a pure bar, line, scatter or other plot.
You can also combine elements, however, for example adding a regression line through a scatter plot.

For example, consider the scatter plot we created earlier:


```{r multiple}
library(tidyverse)
demographics <- read_csv('https://raw.githubusercontent.com/vanatteveldt/ccslearnr/master/data/dutch_demographics.csv')
ggplot(demographics, aes(x = v57_density, y = c_65plus)) +
  geom_point()
```

We can add a regression (trend) line to this plot by adding a `geom_smooth` (using `lm`, i.e. a linear model), which conveniently uses the same aesthetics as the scatter plot:

```{r multiple2}
library(tidyverse)
demographics <- read_csv('https://raw.githubusercontent.com/vanatteveldt/ccslearnr/master/data/dutch_demographics.csv')
ggplot(demographics, aes(x = v57_density, y = c_65plus)) +
  geom_point() + 
  geom_smooth(method = lm)
```

When adding multiple geoms, by default they use the data and aesthetics from the `ggplot` call.
However, you can override these by specifying which data or aesthetics to use for a geom.

For example, we could add the names of selected municipalities, for example the outliers.
To do this, we would add a `geom_text()` with a `label` aesthetic. 
However, if we would label all 345 points it would quickly become a mess.
For that reason, we can create a subset of the data, and use that as the data for the `geom_text`:

```{r multiple3}
subset = filter(demographics, c_65plus > 30 | v57_density > 4000)
ggplot(demographics, aes(x = v57_density, y = c_65plus)) +
  geom_point() + 
  geom_text(data=subset, aes(label=gemeente), nudge_y = -.5)
```

**Solution:**

```{r multiple3-solution}
subset = filter(demographics, c_65plus > 30 | v57_density > 4000 | c_65plus < 15)
ggplot(demographics, aes(x = v57_density, y = c_65plus)) +
  geom_point() + 
  geom_text(data=subset, aes(label=gemeente), nudge_y = -.5)
```



Some things to note here:
- ggplot wants the data as first argument, and the mapping as second. For geoms this is unfortunately the other way around,
  so we specify that the first argument is the data by using `data=subset` rather than just `geom_text(subset,...)`
- `geom_text` now has it's own data, and we also specify a label aesthetic. The x and y aesthetic (i.e. the label positions)
  are not overridden, so they use the mapping from the `ggplot` call. 
- Finally, we use `nudge_y` to move all labels a little bit down so they don't overlap with the points. 

**Exercise:** Can you add labels for the municipalities where less than 15% of the population is over 65?
(hint: you can add another condition to the existing filter call by adding another `|`)

## Creating high-quality graphs for reports

So far, we have mostly kept the default options for graphs and did not spend too much attention to making them nicer to look at or easier to read.
This is fine if you're visualising data as part of your own analysis.
If you want to include the visualizations in a report or article, however, or present them to an audience,
it is worth spending a bit of time to make them both prettier and more informative.
Fortunately, `ggplot` gives you almost endless options to tweak any aspect of each plot.

There are probably as many ideas on the 'best' visualization as there are data scientists,
and many aspects of what makes a graph beautiful or informative are subjective.
However, there are e number of things that you should generally pay attention to:

- The graph should be understandable without reading any text in the rest of the paper or report.
  This means that titles and axis labels should immediately make it clear what the graph is about.
  As a rule of thumb, mention the most important relation or question you are answering in the title,
  and make sure that the axis labels do not contain strange abbreviations or technical terms. 
- Make sure all elements are clearly readable and elements don't overlap each other if it can be helped.
- Make sure the scales and legends are clear and make sense in terms of your data. 

As an example, consider the first graph from the scatter plot section:

```{r ugly}
library(tidyverse)
demographics <- read_csv('https://raw.githubusercontent.com/vanatteveldt/ccslearnr/master/data/dutch_demographics.csv')
ggplot(demographics,
       aes(x = v57_density, y = c_65plus, size=v01_pop, color=v132_income)) +
  geom_point(alpha=.6)
```

This graph is fairly informative: we can see that older people mostly live in smaller, more rural, and somewhat richer communities.
However, if you were to show this graph to an outsider, they would not be able to understand most of the this.
It lacks a descriptive title, uses obscure variable names for axis labels, uses scientific notation for the population,
and personally I find the default style a bit jarring.

These points can all be addressed by **adding extra elements** to the plot. Specifically:

- You can add an overall `ggtitle` and horizontal and vertical axis labels (`xlab` and `ylab`)
- You can alter the way values are mapped between a column and the aesthetic element by adding the `scale_` functions.
  These functions are named for the aesthetic and optionally the data type, so for example `scale_size` changes the size
  aesthetic (for example changing the legend labels and name), while the `scale_color_gradient` creates a gradient scale
  for the color, where you can specify the colors on both ends of the gradient.
- Finally, you can select a `theme`  and then add more `theme(..)` functions to change that theme, for example to 
  change lines, fonts, etc. 
  
The code below addresses most of these points, and as a bonus adds labeling for some of the more interesting points
by first creating a subset of points to label, and then adding a second geom (`geom_text`) using its own data and aesthetics
(see the section on combining geoms).

```{r pretty}
library(tidyverse)
demographics <- read_csv('https://raw.githubusercontent.com/vanatteveldt/ccslearnr/master/data/dutch_demographics.csv')
highlight <- filter(demographics, v01_pop > 250000 | 
                      c_65plus > 30 | 
                      c_65plus > 25 & v57_density > 3000 |
                      c_65plus < 15)
ggplot(demographics,
       aes(x = v57_density, y = c_65plus, size=v01_pop, color=v132_income)) +
  geom_point(alpha=.6) + 
  xlim(0, 7000) + 
  geom_text(data=highlight, aes(label=gemeente), nudge_y = -1, size=3, color="coral") + 
  ggtitle("Where do older people live?",
          "Selected characteristics of Dutch municipalities") +
  xlab("Population density") + 
  ylab("Percentage of population aged 65+") + 
  scale_size(breaks=c(5000, 50000, 500000), labels=c("5.000", "50.000", "500.000"), name="Population") + 
  scale_color_gradient(low="coral", high="darkblue", name="Household\nIncome") +
  theme_classic() + theme(axis.title = element_text(size=12))

```

Have a look at the code above and see if you understand how the different elements combine to create the plot.
Can you change the highlighting criteria to label different points?

As an exercise, pick any of the earlier graphs from this handout, and see if you can turn it into a 
beautiful and informative visualization!





