---
title: 'Dictionary Approaches: Brand Analysis in R'
author: "Wouter van Atteveldt, Kasper Welbers & Philipp Masur"
date: "2021-11"
output: 
  github_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
## include this at top of your RMarkdown file for pretty output
## make sure to have the printr package installed: install.packages('printr')
knitr::opts_chunk$set(echo = TRUE, results = FALSE, message = FALSE, warning = FALSE, fig.keep = "none")
#library(printr)

```


# Introduction

In this tutorial, we are going to learn how to use existing dictionaries to analyze larger text corpora. Dictionaries are often used to conduct a so-called sentiment analysis (and we will run a sentiment analysis at the end of this tutorial), but they can also be used to characterize brands/companies. 

For example, [Nadeau, Rutter and Lettice](https://doi.org/10.1080/0267257X.2020.1764080) (2020) used the [Brand Personality Dictionary](https://provalisresearch.com/products/content-analysis-software/wordstat-dictionary/brand-personality-dictionary/) to understand the process of attitudinal changes towards a brand in crisis and the brand’s communication around the crisis. They collected data (tweets) from brands’ Twitter platforms and use this to conduct extensive text analyses. 

In this tutorial, we will do something very similar using the same dictionary. However, as we will see, there are a lot of problems with these approaches. Particularly, the validity of such dictionaries is - despite them being used in actual research - often questionable. So overall, the main lessons of today's tutorial is that you should always validate to make sure your results are valid for your task and domain. Depending on resources, using crowd coding and/or machine learning can also be a better option than a purely lexical (dictionary-based) approach. 

In this tutorial, we again are going to use functions from the `tidyverse` and the collection of packages around `quanteda`. 

```{r, eval=T}
library(tidyverse)
library(quanteda)
library(quanteda.textplots)
library(quanteda.textstats)
```

# Getting some data

For this example, we will use the latest ~3200 tweets of four different mobile phone companies (Google, Huawei, Samsung, and Sony). If you want, you can scrape their newest tweets using the package `rtweet` and the following code. Note: You will need a Twitter Account to actually scrape the tweets. 

```{r, eval = F}
library(rtweet)

# Get timelines of major mobile phone companies
tmls1 <- get_timeline(c("Google", "Huawei"), n = 3200)         ## we can only scrape two twitter timelines at a time
tmls2 <- get_timelines(c("SamsungMobile", "Sony"), n = 3200)

# Bind them together
tml <- bind_rows(tmls1, tmls2)  ## merging results from both scraping processes

# Save data frame
write_csv(tml, file = "phone_tweets.csv")
```

For the time being, let's simply use the following data set (which I uploaded on Canvas), which contains the latest tweets of these companies from last Monday. 

```{r}
tweets <- read_csv("phone_tweets.csv")
tweets
```

In a first step, we engage in some simple data wrangling. We first create a unique post id and then select specific variables from the comparatively large data set. As a final, but very important step when we use Twitter data, we remove the `#` from the tweets as R will otherwise not recognize e.g., "#awesome" as the same word as "awesome"

```{r}
tweets <- tweets %>%
  group_by(screen_name) %>%
  mutate(post_id = paste(user_id, 1:n(), sep = "_")) %>%
  select(user_id, post_id, created_at, screen_name, description, text, favorite_count, retweet_count) %>%
  mutate(text = str_remove_all(text, "#")) %>%       ## removing '#' from the tweets
  ungroup
tweets
```

When we look at the data, we see that we have a variable that tells us when the tweet was created (created_at), how created the tweet (screen_name), the content of the tweet (text) and some information about whether it was liked (favorite_count) or retweeted (retweet_count).

Because the format is still a simple data set (or tibble), we can quickly check how much each company tweeted over time:

```{r}
# Tweets over time
tweets %>%
  mutate(date = as.Date(created_at)) %>%
  group_by(screen_name, date) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = date, y = n, color = screen_name)) +
  geom_line() +
  theme_bw() +
  labs(color = "Smartphone Provider", x = "Time", y = "Number of tweets")
```

We can clearly see that the last ~3000 tweets of each company cover vastly different time spans. Furthermore, we see that "SamsungMobile" tweets a lot on particular dates (spikes!), but not as much otherwise. Google, of course, which do not only tweet about their phones, tweets a lot more than the others.  


# Getting a DTM

As we learn in the last practical session, the main primitive for any dictionary analysis is the document-term matrix (DTM). 
For more information on creating a DTM from a vector (column) of text, see again the [tutorial on basic text analysis with quanteda](https://github.com/masurp/VU_CADC/blob/main/tutorials/R_text_basics-of-text-analysis.md). For this tutorial, it is important to make sure that the preprocessing options (especially stemming) match those in the dictionary: if the dictionary entries are not stemmed, but the text is, they will not match. In case of doubt, it's probably best to skip stemming altogether.

For our text analysis, we first need to transform this data set into a corpus format. We can do so with the function `corpus()` which only asks us which variable serves as document id (in our case the newly created "post_id") and which contains the actual text ("text"). Note again that all other variables are added as "docvars"!

```{r}
corp <- corpus(tweets, docid_field = 'post_id', text_field = 'text')
corp
```

Now, we can create a document-term matrix as usual. In a first step, we won't apply any preprocessing steps (other than lowercasing, the default in `dfm()`), because dictionaries tend to have full (i.e. not-stemmed) words, and sometimes contain punctuation such as emoticons.

```{r}
dtm <- corp %>% 
  tokens %>%
  dfm
dtm
```

# Dictionary-based analysis

## Obtaining a dictionary

To do a dictionary-based analysis, we can simply use the `dfm_lookup` method to apply an existing dictionary to a dfm (similar to last session). There are many existing dictionaries that can be downloaded from the Internet.

Here, we are going to use the [Brand Personality Dictionary](https://provalisresearch.com/products/content-analysis-software/wordstat-dictionary/brand-personality-dictionary/) (Opoku, Abratt, and Pitt, 2006) that has been used by Nadeau et al. (2020) for classyifing brands with regard to different latent variables:

- Competence
- Excitement
- Ruggedness
- Sincerity
- Sophistication

As it stored in a zip-file on the website, we can use R to download it, unzip and directly transform it into a quanteda dictionary using the function `dictionary()`

```{r}
download.file("https://provalisresearch.com/Download/BrandPersonality.zip", "BrandPersonality.zip")
unzip("BrandPersonality.zip", 'Brand Personality.cat')
brandPers <- read_dict_wordstat(file="Brand Personality.cat")
brandPers %>%
  as_tibble

load(file = "Brand")
```

As we can see, the dictionary includes extensive (> 120 words per category) for each of the five brand personality dimensions. Note: Some of the words are actually not unigrams (i.e., more than one word such as "award_winning"). In this form, they are more or less useless to try to code text that is tokenized into single words. This is a potential drawback of this particular dictionary. We could try to recode the dictionary or try to tokenize the text differently. For the time being, we will just go ahead with this limitation. 

## Applying a quanteda dictionary

Now that we have a dtm and a dictionary, applying it is relatively simple by using the `dfm_lookup` function.
To use the result in further analysis, we convert to a data frame and change it in to a tibble.
The last step is purely optional, but it makes working with it within tidyverse slightly easier:

```{r}
dtm_brand <- dtm %>% 
  dfm_lookup(brandPers)
dtm_brand

result <- dtm_brand %>%
  convert(to = "data.frame") %>% 
  as_tibble
result
```

We know have a data frame that includes information about how often any of the words related to one of the five brand personality dimensions occurred in the tweet. 

In a final step, we know would probably like to combine this results with the initial data set so that we know which text belonged to which company. 

```{r}
tweets2 <- tweets %>% 
  rename(doc_id = post_id) %>%  ## renaming the idea variable to match the result data frame
  inner_join(result) 
tweets2
```

## Analyzing the results

Usually, coding our tweets manually is not really the end goal of our analysis. Instead, we now can use this information to further analyze our text. A first very simple analysis could be to summarize the amount of times each brand personality dimension was mentioned by the different companies. 

```{r}
plot_data <- tweets2 %>%
  group_by(screen_name) %>%
  summarize(competence = sum(COMPETENCE)/n(),
            excitement = sum(EXCITEMENT)/n(),
            ruggedness = sum(RUGGEDNESS)/n(),
            sincerity = sum(SINCERITY)/n(),
            sophistication = sum(SOPHISTICATION)/n())
plot_data
```

Of course, instead of a table, we could also plot the differences using a barplot:

```{r}
plot_data %>%
  pivot_longer(-screen_name) %>%
  ggplot(aes(x = screen_name, y = value, fill = name)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_brewer(palette = "Set2") +
  theme_replace() +
  labs(x = "Smartphone Company", y = "Number of mentions in tweets", fill = "Brand Personality")
```

And to introduce some statistical approaches and how you can do them in R, here is an analysis of variance (ANOVA) testing the differences in "COMPETENCE" across the four companies. 


```{r}
library(emmeans) # don't forget to install the package first!
m_comp <- aov(COMPETENCE ~ screen_name, tweets2)

# Model summary
summary(m_comp)

# Means and confidence intervals
tab <- emmeans(m_comp, specs = c("screen_name"))
tab

# Pairwise comparisons using Tukey's HSD
pairs(tab, adjust = "tukey")
```


## Validating a dictionary

To get an overall quantitative measure of the validity of a dictionary, 
we should manually code a random sample of documents and compare the coding with the dictionary results. 
This can (and should) be reported in the methods section of a paper using a sentiment dictionary.

You can create a random sample from the original data frame using the sample function:

```{r}
sample_ids <- sample(docnames(dtm), size=50)
```

```{r, eval=F}
## convert quanteda corpus to data.frame
docs <- docvars(corp)
docs$doc_id <- docnames(corp)
docs$text <- as.character(corp)

# Create a csv-file of the sub sample
docs %>% 
  filter(doc_id %in% sample_ids) %>% 
  mutate(manual_competence = "",
         manual_excitement = "",
         manual_ruggedness = "",
         manual_sincerity = "",
         manual_sophistication = "") %>% 
  write_csv("to_code.csv")
```

Then, you can open the result in excel, code the documents by filling in the empty columns, and read the result back in and combine with your results above.

Note that I rename the columns and turn the document identifier into a character column to facilitate matching it:

```{r, eval=F}
validation <- read_csv("to_code.csv") %>% 
  mutate(doc_id = as.character(doc_id)) %>% 
  inner_join(result)
```

In the following code, I simply create a "random" manual coding to demonstrate some of the typical validation steps (we will engage in more meaningful validation practices in later practical sessions). 

```{r}
# Just creates just a column with "random" manual codings
validation <- tibble(doc_id=sample_ids, manual_competence = rep(c(1,1, 1, 0, 0), 10)) %>% 
  inner_join(result)
```

Now let's see if my (admittedly completely random) manual coding matches the competence score created by the dictionary analysis. In a first step, we can simply do a correlation:

```{r}
cor.test(validation$manual_competence, validation$COMPETENCE)
```

We can also get a 'confusion matrix' if we create a nominal value from the variable "COMPETENCE" (1 = coded as "competence", 0 = not coded as "competence", thereby neglecting when a tweets mentions several words of a category)

```{r}
validation <- validation %>%
  mutate(competence = ifelse(COMPETENCE > 1, 1, COMPETENCE))
cm <- table(manual = validation$manual_competence, dictionary = validation$competence)
cm
```

This shows the amount of errors in each category. 
Total accuracy is the sum of the diagonal of this matrix divided by total sample size.

```{r}
sum(diag(cm)) / sum(cm)
```


# Sentiment analysis

In this second part of this tutorial, we will engage in a special case of a dictionary analysis: a so-called sentiment analysis. Some of the most important questions about text have to do with *sentiment* (or *tone*): 
Is a text in general positive or negate? Are actors described as likable and successful? 
Is the economy doing well or poorly? Is an issue framed as good or bad?
Is an actor in favor of or against a certain policy proposal? 

*Caveat*: This method is very successful for some tasks such as deciding whether a review is positive or negative. 
In other cases, however, one should be more careful about assuming dictionary analyses are valid. 
Especially in political communication, sentiment can mean one of multiple things,
and many texts contain multiple statements with opposing sentiment.

For more information and critical perspectives on dictionary based sentiment analysis, 
see e.g. the references below:

 + Soroka, S., Young, L., & Balmas, M. (2015). Bad news or mad news? sentiment scoring of negativity, fear, and anger in news content. The ANNALS of the American Academy of Political and Social Science, 659 (1), 108–121.
 + González-Bailón, S., & Paltoglou, G. (2015). Signals of public opinion in online communication: A comparison of methods and data sources. The ANNALS of the American Academy of Political and Social Science, 659(1), 95-107.
 + Barberá, P., Boydstun, A., Linn, S., McMahon, R., & Nagler, J. (2016, August). Methodological challenges in estimating tone: Application to news coverage of the US economy. In Meeting of the Midwest Political Science Association, Chicago, IL.
 

For easy use, the package `SentimentAnalysis` contains 3 dictionaries: `DictionaryGI` is a general sentiment dictionary based on The General Inquirer, and  `DictionaryHE` and `DictionaryLM` are dictionaries of finance-specific words presented by Henry (2008) and Loughran & McDonald (2011) respectively.
The package `qdapDictionaries` also contains a number of interesting dictionaries, including a list of `positive.words` and `negative.words` from the sentiment dictionary of Hu & Liu (2004), and specific lists for `strong.words`, `weak.words`, `power.words` and `submit.words` from the Harvard IV dictionary.

You can inspect the lists and their origins from each package by loading it and getting the help:

```{r}
library(SentimentAnalysis)
?DictionaryGI
names(DictionaryGI)
head(DictionaryGI$negative, 27)

library(qdapDictionaries)
?weak.words
head(weak.words, 27)
```

You can also download many dictionaries as CSV. For example, the VADER lexicon is specifically made for analysing social media and includes important 
words such as "lol" and "meh" that are not present in most standard dictionaries:

```{r}
url <- "https://raw.githubusercontent.com/cjhutto/vaderSentiment/master/vaderSentiment/vader_lexicon.txt"
vader <- read_delim(url, col_names=c("word","sentiment", "details"),  col_types="cdc",  delim="\t")
vader
```

Which dictionary you should use depends on your task and research question. Sometimes, it may even be best to use several ones and compare their validity. 
Again, make sure that your choices for preprocessing match your analysis strategy:
if you want to use e.g. the emoticons in the VADER dictionary, you should not strip them from your data.

## Creating a quanteda dictionary from a word list

You can apply the dictionaries listed above again using dfm_lookup. 
The dictionaries from `SentimentAnalysis` can be directly turned into a quanteda dictionary:

```{r}
GI_dict <- dictionary(DictionaryGI)
```


## Running the sentiment analysis

To run the actual sentiment analysis, we can use the same dtm that we created earlier and again use the `dfm_lookup()` function. Note that we also produce a length variable (i.e., a variable that contains the length of each tweet) that we will use later to compute a sentiment score. 

```{r}
sentiment <- dtm %>%
  dfm_lookup(GI_dict) %>%
  convert("data.frame") %>%
  as_tibble %>%
  mutate(length = ntoken(dtm))
```


Now, we probably want to compute some sort of overall sentiment score. We can make various choices here, 
but a common one is to subtract the negative count from the positive count and divide by either the total number of words or by the number of sentiment words.
We can also compute a measure of subjectivity to get an idea of how much sentiment is expressed in total:

## Creating sentiment scores

```{r}
sentiment <- sentiment %>% 
  mutate(sentiment1 = (positive - negative) / (positive + negative),
         sentiment2 = (positive - negative) / length,
         subjectivity = (positive + negative) / length)
sentiment
```

As a final step, we can again add this data.frame to our original data frame that contains the tweets and companies

```{r}
tweets3 <- tweets %>% 
  rename(doc_id = post_id) %>%  ## renaming the idea variable to match the result data frame
  inner_join(sentiment) 
tweets3
```

For now, let us simply investigate differences in subjectivity across the four companies.

```{r}
tweets3 %>%
  select(screen_name, subjectivity) %>%
  group_by(screen_name) %>%
  summarize(sub = mean(subjectivity)) %>%
  ggplot(aes(x = screen_name, y = sub)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Company", y = "Subjectivity Score")
```


